---
title: "SDS-315-HW-10"
author: "Neena Varanasi"
date: "2024-04-20"
output:
  pdf_document:
    toc: yes

---
    

Name: Neena Varanasi

EID: nv6558

[GitHub Link]()

```{r, message = FALSE, echo=FALSE, warning=FALSE}
#import libraries
library(tidyverse)
library(kableExtra)
library(moderndive)
library(effectsize)
```

## Problem 1- Redlining

### Question: Is there an association between the number of FAIR policies and the racial/ethnic composision of a ZIP code?

### Approach: 
First, I looked at the scatter plot between minorities and the number of FAIR policies along with a linear line of best fit. After visualizing this relationship, I created a linear model and included a interaction term for the fires per 100 housing units, the number of houses built before WWII, and the median family income. Finally, I computed a confidence interval for the linear model.

### Results:
```{r, echo=FALSE, message = FALSE}
redlining = read_csv("redlining.csv")

ggplot(redlining, aes(x=minority, y=policies)) + 
  geom_point() + 
  ggtitle("Minorities vs. Number of FAIR Policies") +
  geom_smooth(method='lm', color='black') +
  theme_classic()


#adjust for fire, age, and income
lm1 = lm(policies ~ minority + fire + age + income, data=redlining)
kable(confint(lm1, level=0.95))

```
The above scatter plot shows the linear relationship between minorities and FAIR policies. There is a increasing trend in the graph.

The above table shows the confidence interval of the main effect between the number of minorities and number of FAIR policies. The confidence interval is 0.0025442	to 0.0141729, meaning it is statistically significant.


### Conclusion:
There is a slight association between the number of minorities and number of FAIR policies as there is a statistically significant relationship between the two variables.

## Problem 2- Grocery store prices
```{r, echo=FALSE, message = FALSE}
#read the dataset in
groceries = read_csv("groceries.csv")

```

### Part A
What kind of price differences do we see across the different stores?
```{r fig.width=7, fig.height=5, fig.align='center', echo=FALSE, message = FALSE}
groceries_store = groceries %>% 
  group_by(Store) %>% 
  summarize(avg_price = mean(Price))

ggplot(groceries_store, aes(x = reorder(Store, avg_price), y = avg_price)) +  
  geom_bar(stat = "identity", fill = "aquamarine4") +
  coord_flip() +
  xlab("Store") +
  ylab("Average Price ($)") +
  ggtitle("Average Grocery Prices Across Stores") +
  theme_classic()
```
The above bar plot shows the average price by store in Texas. The store with the highest price was Whole Foods and the store with the lowest price was Fiesta.

### Part B
```{r, echo=FALSE, message = FALSE}
groceries_product = groceries %>%
  group_by(Product) %>%
  count()

ggplot(groceries_product, aes(x = reorder(Product, n), y = n)) +  
  geom_bar(stat = "identity", fill = "darkseagreen4") +
  coord_flip() +
  xlab("Product") +
  ylab("Number of Stores Selling") +
  ggtitle("Number of Stores Selling Each Product") +
  theme_classic() +
  theme(axis.text.y=element_text(size=7))

```
The above bar plot shows the number of stores selling each product. The only two products that were sold in each store was a Horizon 2% Milk Carton and a carton of eggs. Many products were sold in less than 5 of the 16 stores analyzed.

### Part C

```{r, echo=FALSE, message = FALSE}
lm_grocery = lm(Price ~ Type + Product, data=groceries)

table_grocery = (get_regression_table(lm_grocery, conf.level = 0.95, digits=2))
kable(table_grocery[1:5, ])


```
Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), convenience stores charge somewhere between 0.41 and 0.92 dollars more for the same product.

### Part D
Now fit a model for Price versus Product and Store. Which two stores seem to charge the lowest
prices when comparing the same product? Which two stores seem to charge the highest prices when comparing the same product?
```{r, echo=FALSE, message = FALSE}
lm_store = lm(Price ~ Store + Product, data=groceries)
table_store = (get_regression_table(lm_store, conf.level = 0.95, digits=2))
kable(table_store[1:14, ])

```
Kroger Fresh Fare and HEB charge the lowest prices when comparing the same priduct as seen by the estimates -0.90 and -0.65 respectively.
Whole Foods and Wheatsville Food Co-Op charge the highest prices when comparing the same product as seen by the estimates 0.36 and 0.29 respectively.


### Part E
Central Market charges more for the same product than HEB. Central Market has an estimate of -0.57 and HEB has an estimate of -0.90 when holding the product constant. Therefore, Central Market charges $0.33 more than HEB for the same product on average.

### Part F

```{r, echo=FALSE, message = FALSE}
groceries = mutate(groceries, Income10K = Income/ 10000)
lm_income = lm(Price ~ Income10K + Product, data=groceries)

table_income = (get_regression_table(lm_income, conf.level = 0.95, digits=2))
kable(table_income[1:2, ])


sd_income = standardize_parameters((lm_income), digits=2)
kable(sd_income[1:2, ])

```
Based on the regression table above, we can say that people with a higher Income10K spend 0.005 to 0.03 less on groceries. Therefore, consumers in poorer ZIP codes seem to pay more for the same product on average

The estimated size of the effect of Income10K on Price is -0.03.
